---
title: "Covid91 vaccine study Final2023"
author: "Marbella Sanchez"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "July 24,2023" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="final.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
```

# Introduction

A new preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

$H_0$: The Covid vaccine (drug) is not effective in preventing Covid among subgraoups.
$H_A$: The Covid vaccine (drug) is effective in preventing Covid among subgraups. 


## Males

## Methods 

This statistical analysis we will be using is catergorical ~ categorical, because the COVID vaccine should be effective among subgroups

The method we will be using to display the data us a bar chart, cross table, and chisq test

```{r}
maleData = filter (FinalData, sex =="M")
```


## Results

## Descriptive Results 

### Graphical Descriptive Results 
```{r}
barchartGC(~infected+treatment,data=maleData)
barchartGC(~infected+treatment,data=maleData,type="percent")
```


### Numerical Descriptive Results 
```{r}
table1 <-xtabs(~infected+treatment,data=maleData)
rowPerc (table1)
colPerc (table1)
```
## Inferential Results 
```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```


## Females

## Methods

## Results 
```{r}
femaleData = filter(FinalData, sex == "F")
```


## Descriptive Results 

### Graphical Decriptive Results 

### Numerical Descriptive Results 
```{r}
table2 <- xtabs(~infected + treatment, data=femaleData)
rowPerc(table2)
colPerc(table2)
```
## Inferential Results 
```{r}
chisq.test(table2)
chisqtestGC(table2)
fisher.test(table2)
```
## LGBTQ

## Methods 

## Results 
```{r}
LGBTQData = filter(FinalData, LGBTQ == "gay")
```

## Descriptive Results 

### Graphical Decriptive Results 
```{r}
barchartGC(~infected+treatment,data=LGBTQData)
barchartGC(~infected+treatment,data=LGBTQData, type="percent")
```

### Numerical Descriptive Results 
```{r}
table3 <- xtabs(~infected +treatment, data=LGBTQData)
rowPerc(table3)
colPerc(table3)
```
## Inferential Results 
```{r}
chisq.test(table3)
chisqtestGC(table3)
fisher.test(table3)
```

## Druggies

## Methods 

## Results 
```{r}
DruggiesData = filter(FinalData, DrugUser == "yes")
```


## Descriptive Results

### Graphical Descriptive Results 
```{r}
barchartGC(~infected+treatment,data=DruggiesData)
barchartGC(~infected+treatment,data=DruggiesData, type="percent")
```


### Numerical Descriptive Results

```{r}
table4 <- xtabs (~infected + treatment, data=DruggiesData)
rowPerc(table4)
colPerc(table4)   
```


## Inferential Results 
```{r}
chisq.test(table4)
chisqtestGC(table4)
fisher.test(table4)
```

# Overall Results and Conclusions